---
title: "Statistical analysis"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Load R libraries
```{r results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(brms)
library(bayestestR)
library(kableExtra)
library(ggbeeswarm)
library(RColorBrewer)
library(showtext)
library(lme4)
library(lmerTest)
library(gridExtra)
library(tidybayes)

font_add_google(name = "Lato", family = "Lato", regular.wt = 400, bold.wt = 700)
showtext_auto()
options(stringsAsFactors = FALSE)

SE <- function(x) sd(x) / sqrt(length(x))

get_fixed_effects_with_p_values <- function(brms_model){
  fixed_effects <- data.frame(summary(brms_model)$fixed) %>%
    rownames_to_column("Parameter")
  fixed_effects$p <- (100 - as.data.frame(bayestestR::p_direction(brms_model))$pd) / 100
  fixed_effects %>% select(Parameter, everything()) %>%
    rename(`l-95% CI` = l.95..CI, `u-95% CI` = u.95..CI) %>%
    mutate(` ` = ifelse(p < 0.05, "\\*", " "))
}
```

## Load the data
```{r message=FALSE, warning=FALSE}
fitness_data <- read_csv("data/SR_fitness_data.csv") %>% 
  filter(!is.na(genotype)) %>%
  rename(body_size = `Body size`,
         female_age = `F age`) %>%
  mutate(genotype = factor(genotype, levels = c("STST", "SRST", "SRSR")))

sex_ratio_data <- read.csv("data/SR_sex_ratio_data.csv", stringsAsFactors = FALSE)
sex_ratio_data$n <- with(sex_ratio_data, male + female)
```


## Make a table of summary statistics and sample sizes

Here, we calculate the mean offspring produced by females from each of the three genotypes (STST, SRST, and SRSR), either within each isoline or across all the isolines. We also calculate the % females that failed to produce any offspring, and provide sample size information.
```{r}
means_by_isoline <- fitness_data %>%
  group_by(genotype, Isoline) %>%
  summarise(
    Number_of_females_measured = n(),
    Mean_offspring_per_female = mean(offspring),
    SE = SE(offspring),
    n_females_producing_offspring = sum(offspring != 0),
    Percent_females_producing_offspring = 100 * n_females_producing_offspring / n()) 

means <- fitness_data %>%
  mutate(Isoline = "Across all isolines") %>%
  group_by(genotype, Isoline) %>%
  summarise(
    Number_of_females_measured = n(),
    Mean_offspring_per_female = mean(offspring),
    SE = SE(offspring),
    n_females_producing_offspring = sum(offspring != 0),
    Percent_females_producing_offspring = 100 * n_females_producing_offspring / n()) 

sample_size_table <- bind_rows(means_by_isoline, means) %>%
  rename_all(function(x) gsub("_", " ", x)) %>%
  rename_all(function(x) gsub("Percent", "%", x)) %>%
  rename(Genotype = genotype) 

for_export <- sample_size_table
names(for_export) <- c("Genotype", "Isoline", "n females",
                       "Mean productivity", "SE", "n productive", "% productive")

for_export %>% write_rds("output/sample_size_table.rds")

sample_size_table %>%
  kable(digits = 2) %>% kable_styling()
```

## Fit a model to the data

### Run the Bayesian hurdle model

The model assumes that the response variable, offspring number, is the result of a 'hurdle' process. Essentially this means that the model consists of two sub-models: one controlling the probability that offspring number is non-zero, and one controlling the number of offspring produced provided that more than zero are produced (we assume that offspring number follows a negative binomial distribution, because this improved model fit relative to the simpler hurdle-Poisson model).

We assume that the parameters controlling both the hurdle and the distribution of non-zero values are affected by four fixed effects (the female's genotype: STST, SRST, or SRSR), her isoline, the female's age, and the interaction between genotype and isoline. We also fit two random effects: isoline, and experimental block. All fixed effects were assumed to have a prior distribution following a normal distribution with mean 0 and SD = 5. 

```{r load_all}
if(!file.exists("output/brms_model.rds")){
 
  # The hurdle and the mean have the same set of predictors
  model_formula <- bf(
    offspring ~ genotype * Isoline + female_age + (1 | Block), 
    hu        ~ genotype * Isoline + female_age + (1 | Block)  
  )
  
  model_formula2 <- bf(
    offspring ~ genotype + Isoline + female_age + (1 | Block), 
    hu        ~ genotype + Isoline + female_age + (1 | Block)  
  )
    
  model_formula3 <- bf(
    offspring ~ genotype + female_age + (1 | Block), 
    hu        ~ genotype + female_age + (1 | Block)  
  )
  
  # This model added in response to a reviewer request. 
  # I don't think this model is much use here, since body size is a mediator variable;
  # see the diagram at the top of this page: https://en.wikipedia.org/wiki/Mediation_(statistics)
  # Moreover, many females have no body size measurement, so the sample size is smaller
  body_size_model <- bf(
    offspring ~ genotype + female_age + body_size + (1 | Block), 
    hu        ~ genotype + female_age + body_size + (1 | Block)  
  )
  
  
  # Find R^2 for a brms model, and its 95% CIs, and present neatly
  neat_R2 <- function(model){
    R2 <- bayes_R2(model) %>% round(2)
    paste(R2[1,1], " (95% CIs = ", R2[1,3], "-", R2[1,4], ")", sep = "")
  }
  
  # We set conservative, "regularising" priors - see McElreath's "Statistical Rethinking" textbook
  model_prior <- c(set_prior("normal(0, 3)", class = "b"),
                   set_prior("normal(0, 3)", class = "b", dpar = "hu"))
  
  full_model <- brm(model_formula,
                    family = "hurdle_negbinomial",
                    chains = 4, cores = 1, iter = 40000, inits = 0, seed = 12345,
                    control = list(adapt_delta = 0.9999, max_treedepth = 15),
                    save_all_pars = TRUE, 
                    prior = model_prior, 
                    data = fitness_data)
  
  no_interaction <- brm(model_formula2,
                        family = "hurdle_negbinomial",
                        chains = 4, cores = 1, iter = 40000, inits = 0, seed = 12345,
                        control = list(adapt_delta = 0.9999, max_treedepth = 15),
                        save_all_pars = TRUE, 
                        prior = model_prior, 
                        data = fitness_data)
  
  genotype_only_model <- brm(model_formula3,
                             family = "hurdle_negbinomial",
                             chains = 4, cores = 1, iter = 40000, inits = 0, seed = 12345,
                             control = list(adapt_delta = 0.9999, max_treedepth = 15),
                             save_all_pars = TRUE, 
                             prior = model_prior, 
                             data = fitness_data)
  
  # Added after peer review:
  body_size_model <- brm(body_size_model,
                             family = "hurdle_negbinomial",
                             chains = 4, cores = 1, iter = 40000, inits = 0, seed = 12345,
                             control = list(adapt_delta = 0.9999, max_treedepth = 15),
                             save_all_pars = TRUE, 
                             prior = model_prior, 
                             data = fitness_data)
  
  saveRDS(post_prob(full_model, no_interaction, genotype_only_model), 
          file = "output/model_comparison.rds")
  saveRDS(full_model, file = "output/full_model.rds")
  saveRDS(genotype_only_model, file = "output/genotype_only_model.rds")
  saveRDS(body_size_model, file = "output/body_size_model.rds")
  saveRDS(neat_R2(full_model), file = "output/R2_of_full_model.rds")
  saveRDS(neat_R2(genotype_only_model), file = "output/R2_of_genotype_only_model.rds")
  saveRDS(neat_R2(body_size_model), file = "output/R2_of_body_size_model.rds")
} else{
  full_model <- readRDS("output/full_model.rds")
  genotype_only_model <- readRDS("output/genotype_only_model.rds")
  model_probabilities <- readRDS("output/model_comparison.rds")
  body_size_model <- readRDS("output/body_size_model.rds")
}
```

### Graphically verify the fit of the model using a posterior predictive check

The idea behind posterior predictive checking is that if our model is a good fit, then we should be able to use it to generate a dataset which looks a lot like the dataset we actually observed. Here, we see 11 draws from the 'posterior predictive distribution' (pale blue), which indeed look quite similar to the distribution of the real data (dark blue), suggesting that our model is a good enough approximation of the true data-generating process for reliable inference.

```{r}
pp_check(genotype_only_model, type = "hist", nsamples = 11, binwidth = 5)
```

### Inspect the parameter estimates {.tabset}

```{r}
make_model_table <- function(model){ # helper function for the 3 tables here
  random <- as.data.frame(summary(model)$random[[1]]) %>%
    rownames_to_column("Parameter") %>%
    mutate(p = NA,
           Parameter = c("sd(Block - Intercept)", "sd(Block - Hurdle intercept)"),
           ` ` = "")
  
  rbind(get_fixed_effects_with_p_values(model), random) %>%
    mutate(Parameter = gsub("hu_", "Hurdle - ", Parameter),
           Estimate =  format(round(Estimate, 3), nsmall = 3),
           Est.Error =  format(round(Est.Error, 3), nsmall = 3),
           ` ` = ifelse(p < 0.05, "*", ""),
           ` ` = replace(` `, is.na(` `), ""),
           p = format(round(p, 4), nsmall = 4),
           Rhat = format(round(Rhat, 3), nsmall = 3),
           `l-95% CI` = format(round(`l-95% CI`, 3), nsmall = 3),
           `u-95% CI` = format(round(`u-95% CI`, 3), nsmall = 3),
           Bulk_ESS = round(Bulk_ESS, 0),
           Tail_ESS = round(Tail_ESS, 0)
    ) 
}
```


#### Genotype-only model

This model contains the fixed factor genotype, and the random effect block. 

The response variable (number of progeny) was treated as a hurdle process, i.e. the model estimates the parameters for the probability of producing at least some progeny (the 'Hurdle' parameters), and the number of progeny produced assuming that at least some are (other parameters).

```{r}
no_isoline_model_table <- make_model_table(genotype_only_model)
saveRDS(no_isoline_model_table, "output/no_isoline_model_table.rds")
no_isoline_model_table %>% 
  kable() %>% kable_styling()
```

#### Genotype-by-isoline model

This model contains the fixed factor genotype, the fixed factor isoline, and their interaction, as well as the random effect block. 

The response variable (number of progeny) was treated as a hurdle process, i.e. the model estimates the parameters for the probability of producing at least some progeny (the 'Hurdle' parameters), and the number of progeny produced assuming that at least some are (other parameters).

```{r}
full_model_table <- make_model_table(full_model)
saveRDS(full_model_table, "output/full_model_table.rds")
full_model_table %>% 
  kable() %>% kable_styling()
```

#### Model including body size

This is the same as the top model (i.e. the Genotype-only model), except that it also includes body size a predictor in both components of the hurdle model. Note also that this analysis is restricted to females for which a body size measurement was available, reducing the sample size from 440 to 338 females.

```{r}
bodysize_model_table <- make_model_table(body_size_model)
saveRDS(bodysize_model_table, "output/bodysize_model_table.rds")
bodysize_model_table %>% 
  kable() %>% kable_styling()
```


## Use the model to generate posterior estimates of group means

### Generate posterior predictions of the group means

Here, we estimate the mean for three measures of female fitness using the model, for each genotype (across all isolines) and for each genotype-isoline combination. The model adjusts for variation due to experimental block and female age. 

```{r make_figures, message=FALSE, warning=FALSE}
make_figure_data <- function(by_isoline = FALSE){
  if(by_isoline){
    new <- fitness_data %>% 
      select(genotype, Isoline, body_size, female_age) %>%
      mutate(body_size  = mean(body_size, na.rm = TRUE),
             female_age = mean(female_age)) %>% 
      distinct()
    model <- full_model
    col_names <- paste(new$genotype, new$Isoline, sep = "~")
  } else {
    new <- fitness_data %>% 
      select(genotype, body_size, female_age) %>%
      mutate(body_size  = mean(body_size, na.rm = TRUE),
             female_age = mean(female_age)) %>% 
      distinct()
    model <- genotype_only_model
    col_names <- new$genotype
  }
  
  # Summarise the posterior (dots and CIs in Figure 1 or S1)
  predicted_mean <- data.frame(new, fitted(model, newdata = new, re_formula = NA)) %>% 
    mutate(facet = "A. Mean offspring production")
  predicted_mean_when_fertile <- data.frame(new, fitted(model, newdata = new, dpar = "mu", re_formula = NA)) %>%
    mutate(facet = "B. Mean offspring production\n(excluding infertile females)")
  predicted_prop_fertile <- data.frame(new, fitted(model, newdata = new, dpar = "hu", re_formula = NA)) %>% 
    mutate(facet = "C. % fertile females",
           Estimate = 100 * (1 - Estimate), # Convert to percentage of fertile females, instead of *proportion* that are *in*fertile
           Q2.5 = (1 - Q2.5) * 100, 
           Q97.5 = (1 - Q97.5) * 100)
  
  summary_df <- bind_rows(predicted_mean,
                          predicted_mean_when_fertile,
                          predicted_prop_fertile) %>%
    mutate(genotype = factor(genotype, levels = c("STST", "SRST", "SRSR")))
  if(!by_isoline) summary_df <- summary_df %>% mutate(Isoline = "All isolines")
  
  # Posterior for facet A (overal progeny)
  posterior_means <- fitted(model, newdata = new, re_formula = NA, summary = FALSE) %>% as.data.frame()
  names(posterior_means) <- col_names
  posterior_facetA <- gather(posterior_means) %>% 
    mutate(facet = "A. Mean offspring production")
  
  # Posterior for facet B (excluding infertile females)
  posterior_means <- fitted(model, newdata = new, dpar = "mu", re_formula = NA, summary = FALSE) %>% as.data.frame()
  names(posterior_means) <- col_names
  posterior_facetB <- gather(posterior_means) %>% 
    mutate(facet = "B. Mean offspring production\n(excluding infertile females)")
  
  # Posterior for facet C (% infertile females)
  posterior_means <- fitted(model, newdata = new, dpar = "hu", re_formula = NA, summary = FALSE) %>% as.data.frame()
  names(posterior_means) <- col_names
  posterior_facetC <- gather(posterior_means) %>% 
    mutate(facet = "C. % fertile females")
  
  posterior_df <- bind_rows(
    posterior_facetA, posterior_facetB, posterior_facetC
  )
  
  if(by_isoline){
    posterior_df <- posterior_df %>%
      mutate(split = strsplit(key, split = "~"), 
             genotype = map_chr(split, ~ .x[1]),
             Isoline = map_chr(split, ~ .x[2])) %>% select(-key)
  } else {
    posterior_df <- posterior_df %>%
      rename(genotype = key)
  }
  
  posterior_df <- posterior_df %>% mutate(genotype = factor(genotype, levels = c("STST", "SRST", "SRSR")))
  
  list(summary_df, posterior_df)
}

figure1_data <- make_figure_data()
figureS1_data <- make_figure_data(by_isoline = TRUE)
```

### Plot the posterior predictions of the group means

```{r}
beeswarm_points <- bind_rows(
  fitness_data %>% mutate(facet = "A. Mean offspring production"),
  fitness_data %>% filter(offspring != 0) %>% mutate(facet = "B. Mean offspring production\n(excluding infertile females)")) %>% 
  mutate(Fertility = ifelse(offspring == 0, "Sterile", "Fertile"),
         genotype  = factor(genotype, levels = c("STST", "SRST", "SRSR"))) %>%
    rename(Estimate = offspring) 
  
pal <- c("#6ca0dc", "#e34132")

figure_1 <- figure1_data[[1]] %>%
  ggplot(aes(genotype, Estimate)) + 
  geom_quasirandom(data = beeswarm_points, aes(colour = Fertility),
                size = .7, alpha = 0.6) + 
  geom_errorbar(aes(ymin = Q2.5, ymax = Q97.5), colour = "grey20", size = .8, width = 0) + 
  geom_point(size = 3.1, pch = 21, colour = "black", fill = "grey20") + 
  scale_colour_manual(values = pal) + 
  facet_wrap(~facet, scale = "free_y") + 
  labs(y = "Posterior estimate \u00B1 95% CIs", x = "Genotype") + 
  theme_bw() + 
  theme(strip.background = element_blank(),
        text = element_text(family = "Lato", size = 12),
        panel.grid.major.x = element_blank(), 
        strip.text = element_text(hjust = 0))

figure_S1 <- figureS1_data[[1]] %>%
  ggplot(aes(genotype, Estimate, fill = Isoline)) + 
  geom_errorbar(aes(ymin = Q2.5, ymax = Q97.5), size = .7, width = 0, colour = "grey40", position = position_dodge(0.7)) + 
  geom_point(size = 3.1, pch = 21, colour = "black", position = position_dodge(0.7)) + 
  facet_wrap(~facet, scale = "free_y") + 
  scale_fill_brewer(palette = "Pastel1") +
  labs(y = "Posterior estimate \u00B1 95% CIs", x = "Genotype") + 
  theme_bw() + 
  theme(strip.background = element_blank(),
        text = element_text(family = "Lato", size = 12),
        panel.grid.major.x = element_blank(), 
        strip.text = element_text(hjust = 0))

figure_1 %>% ggsave(filename = "figures/figure_1.pdf", width = 9, height = 4)
figure_S1 %>% ggsave(filename = "figures/figure_S1.pdf", width = 9, height = 4)
```

```{r fig.showtext=TRUE, fig.width = 8, fig.height = 4.5}
figure_1
```
<br></br>
**Figure 1**: The black points and error bars show the posterior estimates of the genotype means for A) offspring production, B) offspring production among the set of females that produced at least one offspring, and C) the percentage of females that produced offspring. The estimates are all derived from a single hurdle model which adjusts for variation due to female age and experimental block, and each estimate is the average across the four isolines (see Figure S1 for estimates split by isoline). The points show the raw values of offspring production for individual females, and are coloured purple for females that produced no offspring. The error bars show the 95% credible intervals on each estimate. 

```{r fig.showtext=TRUE, fig.width = 8.7, fig.height = 4.5}
figure_S1
```
<br></br>
**Figure S1**: The same information as in Figure 1, except split by isoline.



## Calculate pairwise differences between genotypes

**Table 1**: Pairwise comparisons of genotypes for the three measures of female fitness shown in Figure 1: mean offspring production, mean offspring production among females that produced at least one offspring, and the % females that produced at least one offspring. The 'Difference in means' column shows the posterior estimate of the difference between the genotype means, in the original units (i.e. offspring number, or percentage points). A negative difference indicates that the genotype with more copies of SR has lower female fitness, the parentheses show the error and 95% quantiles of the posterior difference in means. The ‘Relative difference’ column expresses each difference in relative terms; e.g. the first row shows that the mean number of offspring produced by SR/ST females was 92% as much as the number produced by ST/ST females, with 95% confidence limits of 70-110%. Finally, $p$ is the posterior probability that the true difference in means is zero or of the opposite sign to the estimate shown here (similar to a conventional $p$-value).

```{r}
compare_means <- function(mean1, mean2, posterior){
  
  posterior <- posterior %>%
    filter(genotype %in% c(mean1, mean2)) %>% 
    select(genotype, value) %>% mutate(draw = rep(1:(n() / 2), 2)) %>%
    spread(genotype, value)
  
  abs_difference <- as_tibble(posterior_summary(as.mcmc(posterior[, mean2] - posterior[, mean1])))
  rel_diff <- as_tibble(posterior_summary(as.mcmc(posterior[, mean2] / posterior[, mean1])))
  p_value <- as.numeric(100 - p_direction(posterior[, mean2] - posterior[, mean1])) / 100
  
  tibble(
    Comparison = paste(mean1, mean2, sep = " \u2192 "),
    `Fitness trait` = NA,
    `95% CIs abs` = paste(" (", format(round(abs_difference$Q2.5, 1), nsmall = 1), " to ", format(round(abs_difference$Q97.5, 1), nsmall = 1), ")", sep = ""),
    `Difference in means` = paste(format(round(abs_difference$Estimate, 2), nsmall = 2), `95% CIs abs`, sep = ""),
    Error1 = abs_difference$Est.Error,
    `95% CIs rel` = paste(" (", format(round(rel_diff$Q2.5, 1), nsmall = 1), " to ", format(round(rel_diff$Q97.5, 1), nsmall = 1), ")", sep = ""),
    `Relative difference` = paste(format(round(rel_diff$Estimate, 2), nsmall = 2), `95% CIs rel`, sep = ""),
    Error2 = rel_diff$Est.Error,
    p = p_value
  ) %>% select( -`95% CIs abs`, -`95% CIs rel`)
 
}

table_of_contrasts <- bind_rows(
  compare_means("STST", "SRST", figure1_data[[2]] %>% filter(facet == "A. Mean offspring production")),
  compare_means("STST", "SRSR", figure1_data[[2]] %>% filter(facet == "A. Mean offspring production")),
  compare_means("SRST", "SRSR", figure1_data[[2]] %>% filter(facet == "A. Mean offspring production")),
  compare_means("STST", "SRST", figure1_data[[2]] %>% filter(facet == "B. Mean offspring production\n(excluding infertile females)")),
  compare_means("STST", "SRSR", figure1_data[[2]] %>% filter(facet == "B. Mean offspring production\n(excluding infertile females)")),
  compare_means("SRST", "SRSR", figure1_data[[2]] %>% filter(facet == "B. Mean offspring production\n(excluding infertile females)")),
  compare_means("STST", "SRST", figure1_data[[2]] %>% filter(facet == "C. % fertile females")),
  compare_means("STST", "SRSR", figure1_data[[2]] %>% filter(facet == "C. % fertile females")),
  compare_means("SRST", "SRSR", figure1_data[[2]] %>% filter(facet == "C. % fertile females"))
) %>% mutate(`Fitness trait` = rep(c("Mean offspring production",
                                     "Mean offspring production (excluding infertile females)",
                                     "% fertile females"), each = 3)) %>%
  mutate(Error1 = format(round(Error1, 2), nsmall = 2),
         Error2 = format(round(Error2, 2), nsmall = 2),
         ` ` = ifelse(p < 0.05, "*", " "),
         p = format(round(p, 4), nsmall = 4))

table_of_contrasts %>%
  kable() %>% kable_styling()
```


## Make wing length figure

```{r body_size_fig}
p1 <- fitness_data %>% 
  filter(!is.na(body_size)) %>%
  ggplot(aes(genotype, body_size, colour = genotype)) + 
  geom_quasirandom(alpha = 0.5) + 
  stat_summary(fun.data = "mean_cl_boot", colour = "grey20") +
  theme_bw() + ylab(NULL) + xlab("Genotype") +
  theme(strip.background = element_blank(),
        legend.position = "none",
        panel.grid.major.x = element_blank(), 
        strip.text = element_text(hjust = 0))

p2 <- fitness_data %>% 
  filter(!is.na(body_size)) %>%
  ggplot(aes(body_size, fill = genotype)) + 
    geom_density(alpha = 0.3) + 
    theme_bw() + 
    coord_flip() + xlab(NULL) + ylab("Density") +
    theme(strip.background = element_blank(),
          legend.position = "none",
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          panel.grid.major.x = element_blank(), 
          panel.grid.minor.x = element_blank(), 
          strip.text = element_text(hjust = 0))

grid.arrange(p1, p2,
             widths = c(0.7, 0.32),
             ncol = 2, left = "Wing vein length (mm)"
)
```
<br></br>
**Figure 2:** Distribution of wing lengths for each genotype, showing the individual values (left) or the frequency distribution (right).

```{r echo=FALSE}
ggsave(
  grid.arrange(p1 + theme(text = element_text(family = "Lato", size = 12)), 
               p2 + theme(text = element_text(family = "Lato", size = 12)), 
               widths = c(0.7, 0.32),
               ncol = 2, left = "Wing vein length (mm)"
  ), filename = "figures/figure_2.pdf", height = 3.3, width = 5)
```


### Associated statistics for Figure 2

#### Mody size means
```{r}
body_size_model_data <- fitness_data %>%
  filter(!is.na(body_size)) %>%
  mutate(body_size_scaled = as.numeric(scale(body_size)))

body_size_model_data %>%
  group_by(genotype) %>%
  summarise(`Mean wing vein length (mm)` = mean(body_size),
            SE = sd(body_size) / sqrt(n())) %>%
  kable(digits=3) %>% kable_styling()
```

#### Linear mixed model
Note that body size data is missing for one body size - isoline combination, so we did not fit the interaction term (preventing rank deficiency). The model shows that body size differs significantly among genotypes and isolines.
```{r}
body_size_model <- lmer(body_size_scaled ~ genotype + Isoline + (1 | Block), 
                      data = body_size_model_data)
summary(body_size_model)
anova(body_size_model)
```

#### Sample sizes for the body size data
```{r}
sample_sizes <- body_size_model_data %>%
  group_by(genotype, Isoline) %>%
  summarise(n = n())

overall <- sample_sizes %>%
  group_by(genotype) %>%
  summarise(n =sum(n)) %>%
  mutate(Isoline = "Across all isolines") %>%
  select(!! names(sample_sizes))

sample_sizes %>%
  bind_rows(overall) %>%
  kable(digits=3) %>% kable_styling()
```

## Analysis of offspring sex ratio

```{r}
if(!file.exists("output/SR_model_comparison.rds")){
  sex_ratio_genotype <- brm(
    female | trials(n) ~ genotype,
    data = sex_ratio_data,
    save_all_pars = TRUE,
    family = "binomial",
    chains = 4, iter = 40000, cores = 1)
  
  sex_ratio_isoline <- brm(
    female | trials(n) ~ genotype + isoline,
    data = sex_ratio_data,
    save_all_pars = TRUE,
    family = "binomial",
    chains = 4, iter = 40000, cores = 1)
  
  sex_ratio_interaction <- brm(
    female | trials(n) ~ genotype * isoline,
    data = sex_ratio_data,
    save_all_pars = TRUE,
    family = "binomial",
    chains = 4, iter = 40000, cores = 1)
  
  # Save model tables
  sex_ratio_genotype_model <- get_fixed_effects_with_p_values(sex_ratio_genotype)
  sex_ratio_isoline_model <- get_fixed_effects_with_p_values(sex_ratio_isoline)
  sex_ratio_interaction_model <- get_fixed_effects_with_p_values(sex_ratio_interaction)
  
  # Calculate posterior model probabilities
  SR_model_comparison <- post_prob(sex_ratio_genotype, 
                                   sex_ratio_isoline, 
                                   sex_ratio_interaction)
  
  # Predict sex ratio across all isolines, or within each isoline
  new <- sex_ratio_data %>% 
    select(isoline, genotype) %>% mutate(n = 100)
  preds_all_isolines <- as_tibble(as.data.frame(fitted(sex_ratio_genotype, new[1:3,], summary = FALSE)))
  preds_by_isoline <- as_tibble(as.data.frame(fitted(sex_ratio_interaction, new, summary = FALSE)))
  names(preds_all_isolines) <- new$genotype[1:3]
  names(preds_by_isoline) <- paste(new$genotype, new$isoline, sep = "_")
  
  saveRDS(sex_ratio_genotype_model, file = "output/sex_ratio_genotype_model.rds")
  saveRDS(sex_ratio_isoline_model, file = "output/sex_ratio_isoline_model.rds")
  saveRDS(sex_ratio_interaction_model, file = "output/sex_ratio_interaction_model.rds")
  saveRDS(SR_model_comparison, file = "output/SR_model_comparison.rds")
  saveRDS(preds_all_isolines, file = "output/preds_all_isolines.rds")
  saveRDS(preds_by_isoline, file = "output/preds_by_isoline.rds")
} else {
  sex_ratio_genotype_model <- readRDS("output/sex_ratio_genotype_model.rds")
  sex_ratio_isoline_model <- readRDS("output/sex_ratio_isoline_model.rds")
  sex_ratio_interaction_model <- readRDS("output/sex_ratio_interaction_model.rds")
  SR_model_comparison <- readRDS("output/SR_model_comparison.rds")
  preds_all_isolines <- readRDS("output/preds_all_isolines.rds")
  preds_by_isoline <- readRDS("output/preds_by_isoline.rds")
  sex_ratio_genotype_model <- readRDS("output/sex_ratio_genotype_model.rds")
}
```

### Posterior model probabilities
The top model (with >99% probability) contains genotype and isoline (just the main effects). The model with an interaction, and the model lacking the isoline effect, had much lower probability. This indicated that SR/ST genotype affected the offspring sex ratio, and so did isoline, but there was little/no evidence for an interaction.
```{r}
round(sort(SR_model_comparison, decreasing=T), 3) %>% 
  kable() %>% kable_styling()
```

### Model results {.tabset}

#### Top model (genotype + isoline)
```{r}
sex_ratio_isoline_model %>% kable(digits = 3) %>% kable_styling()
```

#### Genotype x isoline model
```{r}
sex_ratio_interaction_model %>% 
  kable(digits = 3) %>% kable_styling()
```

#### Genotype only model
```{r}
sex_ratio_genotype_model %>% 
  kable(digits = 3) %>% kable_styling()
```

### Estimated offspring sex ratio by genotype
```{r}
preds_all_isolines %>%
  map(~ posterior_summary(.x)) %>% do.call("rbind", .) %>% as_tibble() %>%
  mutate(Genotype = names(preds_all_isolines)) %>%
  select(Genotype, everything()) %>% 
  kable(digits = 3) %>% kable_styling()
```

### Estimated offspring sex ratio by genotype and isoline
```{r}
preds_by_isoline %>%
  map(~ posterior_summary(.x)) %>% do.call("rbind", .) %>% as_tibble() %>%
  mutate(Genotype = sex_ratio_data$genotype,
         Isoline = sex_ratio_data$isoline) %>%
  select(Genotype, Isoline, everything()) %>% 
  kable(digits = 3) %>% kable_styling()
```

### Hypothesis testing
Here, we calculate the posterior difference in the mean sex ratio between each of the 3 genotypes:
```{r}
get_p <- function(x) (100 - as.numeric(p_direction(x)))/100


data.frame(
  Comparison = c("SR/SR - ST/ST", "SR/ST - ST/ST", "SR/SR - SR/ST"),
  rbind(posterior_summary(preds_all_isolines$SRSR - preds_all_isolines$STST),
        posterior_summary(preds_all_isolines$SRST - preds_all_isolines$STST),
        posterior_summary(preds_all_isolines$SRSR - preds_all_isolines$SRST))
) %>% mutate(
  p = c(get_p(preds_all_isolines$SRSR - preds_all_isolines$STST),
        get_p(preds_all_isolines$SRST - preds_all_isolines$STST),
        get_p(preds_all_isolines$SRSR - preds_all_isolines$SRST))
) %>%
  kable(digits = 3) %>% kable_styling()
```

### Figure showing sex ratios
```{r}
SR_figure <- gather(preds_all_isolines) %>%
  ggplot(aes(key, value, fill = key)) + 
  geom_eye() + 
  geom_hline(yintercept = 50, linetype = 2) + 
  coord_cartesian(ylim = c(49, 65)) + 
  theme_bw() +
  theme(legend.position = "none") + 
  ylab("% daughters among adult offspring\n(posterior estimate)") + 
  xlab("Maternal genotype") 

SR_figure
```

```{r echo=FALSE}
(SR_figure + theme(text = element_text(family = "Lato", size = 12))) %>%
  ggsave(filename = "figures/SR_figure.pdf", height = 4.7, width = 5)
```



### Figure showing sex ratios - split by isoline
```{r}
SR_figure2 <- gather(preds_by_isoline) %>%
  mutate(split = strsplit(key, split = "_"),
         Genotype = map_chr(split, ~.x[1]),
         isoline = paste(map_chr(split, ~.x[2]), "isoline")) %>%
  select(value, Genotype, isoline) %>%
  ggplot(aes(Genotype, value, fill = Genotype)) + 
  geom_eye() + 
  geom_hline(yintercept = 50, linetype = 2) + 
  facet_wrap(~isoline) + 
  coord_cartesian(ylim = c(48, 71.3)) + 
  theme_bw() +
  theme(legend.position = "none") + 
  ylab("% daughters among adult offspring\n(posterior estimate)") + 
  xlab("Maternal genotype")

SR_figure2
```

```{r echo=FALSE}
(SR_figure2 + theme(text = element_text(family = "Lato", size = 12))) %>%
  ggsave(filename = "figures/SR_figure_by_isoline.pdf", height = 7.8, width = 8)

```

